---
title: "Data Import"
output: github_document
---
Command option I to create data chunk

Start loading data (reader is pt of tidyverse)
```{r}
library(tidyverse)
```

## Let's import a dataset
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA",".","")) # Fix the missingness
litters_df = 
  janitor::clean_names(litters_df) # cleaning up the names

pups_df = 
  read_csv("data/FAS_pups.csv", skip = 3, na = c("NA",".",""))
pups_df = 
  janitor::clean_names(pups_df)

```

## `select`
select specific variable
```{r}
select(litters_df, group, litter_number, gd0_weight) # after select, gives a data frame with only these variables
```

select a range
```{r}
select(litters_df, group:gd_of_birth)
# keep all variables between this and this
```

select by removal
```{r}
select(litters_df, -group)
select(litters_df, -(group:gd_of_birth))
```

select by prefix, keeping everything that starts with gd, and also keep group
```{r}
select(litters_df, group, starts_with("gd"))
# also we have end with & contain
```

rearrange by selecting
```{r}
select(litters_df, starts_with("gd"), group) #rearrange the order
select(litters_df, litter_number, everything()) # first give litter_number and then everything else

relocate(litters_df, litter_number) # does the same thing with select and everything
```

rename by selecting
```{r}
select(litters_df, GROUP = group, everything()) # rename this variable 

rename(litters_df, GROUP = group)
```


What if i only select one variable...?
```{r}
select(litters_df, group) # give me a data frame with just one variable 

# Select is different with pull
pull(litters_df, group) # gives a vector, or $, almost never use this, keep everything in data frame
litters_df$group
```

Learning Assessment: In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

use `filter()` to filter out rows
```{r}
filter(litters_df, gd_of_birth == 20, pups_born_alive == 16) # test if value = 20, keeps those returned true, others filtered out
```

look for more than 5 pups alive
```{r}
filter(litters_df, pups_born_alive > 5)
filter(litters_df, pups_born_alive >= 5)
filter(litters_df, pups_born_alive <= 5)
```

look for non-matches
```{r}
filter(litters_df, pups_born_alive != 6)
```

match to a set
```{r}
#suppose i want in this dataset
filter(litters_df, group %in% c("Con7",("Con8")))
```

filter out missing rows with `drop_na()`
Drop rows containing missing values
```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)
```



Learning Assessment: In the pups data:

Filter to include only pups with sex 1
Filter to include only pups with PD walk less than 11 and sex 2
Solution
```{r}
filter(pups_df, sex == 1)
filter(pups_df, pd_walk <= 11, sex == 2)
```

## `mutate`
 use `mutate()` to create or modify variables. 
```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group) # update variable from uppercase to lower
)
```
 
## `arrange` 

`arrange()` arranges things
```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, group, pups_born_alive)
arrange(litters_df, desc(pups_born_alive)) # we can put things in descending
arrange(litters_df, desc(group), pups_born_alive) # group in desc while pups in asc

arrange(litters_df, gd0_weight) # turns out arrange makes NA end up at the end
```

## PIPES
 
 Let's see a bad way to do things
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA",".",""))

litters_df = 
  janitor::clean_names(litters_df)

litters_df = 
  select(litters_df, group, litter_number, starts_with("gd"))

litters_df = 
  drop_na(litters_df)

litters_df = 
  mutate(
    litters_df,
    wet_gain = gd18_weight - gd0_weight
  )
```
 
other bad ways: 
nesting things together

Use pipes to avoid all this
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA",".","")) |> # command+shift+m - pipe operator hotkey
  janitor::clean_names() |> 
  select(group, litter_number, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight
  )
```

## hotkey
* **highlight command enter**, only highlight the middle part (without litters_df and the end bracket) so it outputs the result 
* also can put your cursor in the chunk, **command enter**, r is gonna  run the whole chunk
* press up to run the last piece
* or press up to view


Learning Assessment: Write a chain of commands that:

loads the pups data
cleans the variable names
filters the data to include only pups with sex 1
removes the PD ears variable
creates a variable that indicates whether PD pivot is 7 or more days
```{r}
pups_df = (
  read_csv("data/FAS_pups.csv", skip = 3, na = c("NA",".","")) |> 
    janitor::clean_names() |> 
    filter(sex == 1) |> 
    select(-pd_ears) |> 
    mutate(pd_pivot_over_7  = pd_pivot >= 7)
)
```

write a line, run it, write a line, run it

## tidyverse assumes data comes first

If we do:
```{r}
litters_df |> 
  lm(wt_gain ~ gd_of_birth)
```
error: cannot coerce class ‘"formula"’ to a data.frame

```{r}
litters_df |> 
  lm(wt_gain ~ gd_of_birth, data = _)
```

