---
title: "Data Import"
output: github_document
---
Command option I to create data chunk

Start loading data (reader is pt of tidyverse)
```{r}
library(tidyverse)
```

## Let's import a dataset
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA",".","")) # Fix the missingness
litters_df = 
  janitor::clean_names(litters_df) # cleaning up the names

pups_df = 
  read_csv("data/FAS_pups.csv", skip = 3, na = c("NA",".",""))
pups_df = 
  janitor::clean_names(pups_df)

```

## `select`
select specific variable
```{r}
select(litters_df, group, litter_number, gd0_weight) # after select, gives a data frame with only these variables
```

select a range
```{r}
select(litters_df, group:gd_of_birth)
# keep all variables between this and this
```

select by removal
```{r}
select(litters_df, -group)
select(litters_df, -(group:gd_of_birth))
```

select by prefix, keeping everything that starts with gd, and also keep group
```{r}
select(litters_df, group, starts_with("gd"))
# also we have end with & contain
```

rearrange by selecting
```{r}
select(litters_df, starts_with("gd"), group) #rearrange the order
select(litters_df, litter_number, everything()) # first give litter_number and then everything else

relocate(litters_df, litter_number) # does the same thing with select and everything
```

rename by selecting
```{r}
select(litters_df, GROUP = group, everything()) # rename this variable 

rename(litters_df, GROUP = group)
```


What if i only select one variable...?
```{r}
select(litters_df, group) # give me a data frame with just one variable 

# Select is different with pull
pull(litters_df, group) # gives a vector, or $, almost never use this, keep everything in data frame
litters_df$group
```

Learning Assessment: In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

use `filter()` to filter out rows
```{r}
filter(litters_df, gd_of_birth == 20, pups_born_alive == 16) # test if value = 20, keeps those returned true, others filtered out
```

look for more than 5 pups alive
```{r}
filter(litters_df, pups_born_alive > 5)
filter(litters_df, pups_born_alive >= 5)
filter(litters_df, pups_born_alive <= 5)
```

look for non-matches
```{r}
filter(litters_df, pups_born_alive != 6)
```

match to a set
```{r}
#suppose i want in this dataset
filter(litters_df, group %in% c("Con7",("Con8")))
```

filter out missing wows with `drop_na()`
Drop rows containing missing values
```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)

```


Learning Assessment: In the pups data:

Filter to include only pups with sex 1
Filter to include only pups with PD walk less than 11 and sex 2
Solution
```{r}
filter(pups_df, sex == 1)
filter(pups_df, pd_walk <= 11, sex == 2)
```


