---
title: "Tidy data"
output: github_document
---
Command option I to create data chunk

Start loading data (reader is pt of tidyverse)
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Rules for tidy data
* columns are variables
* rows are observations
* every value has a call

## Why tidy your data?
* allow you to manipulate model using this tool that people over the world have created and use
* data written for computers is easy to work with
* takes disparate data source, widow down to the same structure; tidy data are all the same, untidy are different in different/weird ways
* Some data aren't really amenable to tidiness
  - genomics
  - neuro-imaging, single cell coded multiple info

## Important hotkeys
* command+shift+m - pipe operator hotkey
* **highlight command enter**, only highlight the middle part (without litters_df and the end bracket) so it outputs the result 
* also can put your cursor in the chunk, **command enter**, r is gonna  run the whole chunk
* press up to run the last piece
* or press up to view
## Start Coding now!
## Pivit longer

```{r}
pulse_df = # command+shift+m - pipe operator hotkey
  read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    values_to = "bdi_score", # bdi stands for bed depression score, 
    names_prefix = "bdi_score_" # deleting the prefix
  ) |> 
  mutate(visit = replace(visit, visit == "bl", "00m")) |> # replacing baseline to 00m
  relocate(id, visit)
```
View(pulse_df) # to see demographics

## One more example

```{r}

litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA",".","")) |> 
  janitor::clean_names() |> 
  pivot_longer(# find which column we want to make longer
    cols = gd0_weight:gd18_weight, # now each one have two rows, one is gd0 one is gd18 other data are just reproduced
    names_to = "gd_time",
    values_to = "weight"
  ) |> 
  mutate(gd_time = case_match(
    gd_time, 
    "gd0_weight" ~ 0,
    "gd18_weight" ~ 18
  )) #since 18 and 0 are in the middle, cannot use prefix function...
# other functions like case_when also works...


```


## Pivot wider

```{r}

analysis_df= 
  tibble(group= c("treatment", "treatment", "control","control"),
         time = c("pre","post","pre","post"),
         mean = c(4,10,4.2,5)
         )
```

Pivot wider for human readability

```{r}

analysis_df |> 
  pivot_wider( # what are the variables that we want to create names from
    names_from = time,
    values_from = mean
  ) |> 
  knitr::kable() # create output format, a neater text image, more importantly, it produce the output in ur rmd file is a nicely formatted table
```


## Bind tables

```{r}

# when we have tables with similar format, we stack tables on top of each other
# use read_excel, pull out the data, put them in dataframe
# first table: B3toD6

# always helpful using View(fellowship_ring)

fellowship_ring = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") |> 
  mutate(movie = "fellowship_ring") 

two_towers = 
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") |> 
  mutate(movie = "two_towers")

return_king = 
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") |> 
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(fellowship_ring, two_towers, return_king) |> #take as many datasets as you want
  janitor::clean_names() |> 
  pivot_longer(
    cols = female:male,
    names_to = "sex",
    values_to = "words"
  ) |> 
  relocate(movie) |> 
  mutate(race = str_to_lower(race))

```

## Join FAS datasets

Import `litters` dataset

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA",".","")) |> 
  janitor::clean_names() |> 
  mutate(wt_gain = gd18_weight - gd0_weight) |> # non-tidyness due to excel have multiple info, col-name have both group name and group number, con/mod/...
  separate(
    group, into = c("dose","day_of_treatment"), sep = 3 # separate into these, by index
  )
  # if group name have different length, first we need to use case_match(), case_when
  # cuz separate only works when there's a pattern
```

Import `pups` next!

```{r}

pups_df = 
  read_csv("data/FAS_pups.csv", skip = 3, na = c("NA",".","")) |> 
  janitor::clean_names() |> 
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  )
```

Join the datasets!

```{r}
# prioritize pups, keep every entry of this data-frame and add litters data-frame, use left_join

fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") |> # by argument to specify what to join by
  relocate(litter_number, dose, day_of_treatment)

```













